Stack Safety as a Security Property

### 2020-07-08 main.tex
Better presentation of the testing property

In such a system, all memory addresses, registers and the program counter are
enriched with an abstract metadata tag, which can point to the representation of
an arbitrary data structure. A micropolicy is thus defined by a collection of
rules to be checked as each instruction executes and together define the
transfer function of the security policy. Each rule inspects the tags associated
to the inputs of an instruction (PC, registers, memory) as well as the
instruction opcode itself, and produces one of two outcomes: either the
instruction is allowed to execute (and tags for the new PC and its result are
produced and assigned to those), or the machine fail-stops with a policy
violation.

Efficient execution of these micropolicies relies on hardware implementations
such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
maps tag inputs to outputs and a mechanism to trap misses to software handlers
and thereby computes the transfer function of the software-defined policy.
Previous work has shown that a wide range of micropolicies can be defined and
enforced efficiently \rb{cite}.
%
\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
policies}

\paragraph*{A conservative policy}
%
The first micropolicy we present for eager stack safety is very similar to the
Depth Isolation policy presented by Roessler and DeHon \rb{cite}. Their policies
tag the stack memory with ownership information associated to each stack frame
(and to each separate object inside that frames, a more finely grained access
control that we do not need to consider), and registers holding stack pointers
with matching access permission information. Beyond those capabilities, the
compiler generates designated sequences of instructions with special tags that
grant them permission to manage the stack.

Our micropolicy works as follows. We assume for ease of exposition that no
arguments are passed on the stack \rb{Come back to this later?}. Initially, the
entire stack is tagged as unused with $\tagNoDepth$ \rb{actually, the ``initial
  frame'' should be tagged with its proper depth}, and the {\PCname} carries in
its tag at all times the depth of the current activation, starting with
$\tagPCDepth{0}$. The current activation depth needs to change exactly at those
points in the program where the contour of the stack safety property changes:

@@

\item From caller to callee, when the machine executes an instruction tagged as
  belonging to the call map with $\tagCall$. At this point, the {\PCname} depth
  is incremented to, say $\tagPCDepth{n}$; on function entry, the locations of
  the new frame is marked $\tagStackDepth{n}$.

\item From callee back to caller, when execution reaches the instruction
  immediately following the original call instruction and the $\SP$ is the same
  as at the time of that call. At this point, the {\PCname} is decremented;
  previously, the locations of the old frame are restored to $\tagNoDepth$.

@@

These sequences also enforce the $\SP$ discipline. When stack frames are of
fixed size, it suffices to increment $\SP$ and give it with a special tag in the
header sequence, which is later checked after decrementing $\SP$ during the exit
sequence to authorize the return. The well-formedness of the blessed sequences
can be checked statically. Only programs whose sequences are well-formed are
protected by the micropolicy, although programs are still protected (by
fail-stopping) even if the intended sequences are missing. By tracking the
execution of the blessed sequences in the {\PCname} tag state, the micropolicy
ensures that the entry sequence is executed correctly a declared call.

@@

\paragraph*{Testing the conservative policy}
%
This conservative policy is can be shown to satisfy an eager \emph{testing
property}, which enforces lockstep and performs integrity and confidentiality
tests for a stack of running variants at each step.

\definition{Eager integrity test}
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{eager integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
    \]

\definition{Eager confidentiality test}
  Given two machine states $\mach_1$ and $\mach_2$ with stepping states and
  observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$ these satisfy the \emph{eager confidentiality
  test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \obs_1 ~ \mach_1' ~ \mach_2 ~ \obs_2 ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\definition{Variant stacks}
  The testing property keeps track of execution information on a stack whose
  elements record an initial machine, an initial and current variant of the
  machine, a contour and a return map:
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
  Given the current machine state $\mach$, a stack $\vs$ is \emph{well-formed}
  with respect to it if:
  \begin{itemize}
  \item The current variant is indeed a variant of the current state.
  \item The stack is nonempty and the bottom return map is the false property.
  \item The current state is in the trace of every initial machine.
  \item The variant state is in the trace of its own initial variant.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item All contours have public registers.
  \end{itemize}

@@

  \item If the machine fail-stops,
    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
    the property is safisfied for the singleton trace $(\mach, \pol)$ at any
    depth up to the size of $\vs$.

  \item If the top variant of the stack (and only the top variant) returns,
    the property is safisfied for the singleton trace $(\mach, \pol)$ at depth
    equal to the size of the stack. \rb{Avoid confusion with return below}

  \item If the machine takes a step
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    that is neither a call nor a return, all
    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
    (these tests are required also for calls and returns, and omitted in those cases),
    and the stepped $(\mach', \pol')$ and $\vs'$ satisfy the eager testing property
    for a given depth $d$ \rb{omitting stack size condition?} and MP-trace $\MPT$,
    then the property is also satisfied for the extension of the trace \rb{notation}
    at depth $d$.

>>>>

In such a system, all memory addresses and registers (including the program
counter) are enriched with an abstract metadata tag, which can point to the
representation of an arbitrary data structure. A micropolicy is thus defined by
a collection of rules to be checked as each instruction executes and together
define the transfer function of the security policy. Each rule inspects the tags
associated to the inputs of an instruction (PC, registers, memory) as well as
the instruction opcode itself, and produces one of two outcomes: either the
instruction is allowed to execute (and generates and assigns tags for the new PC
and the result of the operation), or the machine fail-stops with a policy
violation.

Efficient execution of these micropolicies relies on hardware implementations
such as the PUMP architecture \rb{cite}, which include a monitor rule cache that
maps tag inputs to outputs and a mechanism to trap misses to software handlers
and thereby compute the transfer function of the software-defined policy.
Previous work has shown that a wide range of micropolicies can be defined and
enforced efficiently \rb{cite}.
%
\rb{Maybe go into more detail, examples later\ldots or try to segue into stack
policies}

\paragraph*{A conservative policy}
%
The first micropolicy we present to enforce the definition of stack safety
developed in \cref{sec:lse-and-wbcf} is very similar to the Depth Isolation
policy presented by \citet{DBLP:conf/sp/RoesslerD18}. Their policy tags the
stack memory with ownership information associated to each stack frame (and to
each separate object inside that frames, a more finely grained access control
that we do not need to consider here), and registers holding stack pointers with
matching access permission information. Beyond those capabilities, the compiler
generates designated sequences of instructions with special tags that grant them
permission to manage the stack, say during calls and returns. \rb{Conceivably on
alloc operations, etc., which we have not yet mentioned.}

Our micropolicy works as follows. We assume for ease of exposition that no
arguments are passed on the stack \rb{Come back to this later?}. Initially, the
entire stack is tagged as unused with an $\tagNoDepth$ tag \rb{actually, the
``initial frame'' should be tagged with its proper depth}, and the {\PCname}
carries in its tag at all times the depth of the current activation, starting
with a $\tagPCDepth{0}$ tag. The current activation depth needs to change
exactly at those points in the program where the contour of the stack safety
property changes:

@@

\item From caller to callee, when the machine executes an instruction tagged
  with $\tagCall$ as belonging to the call map. At this point, the current
  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
  the locations of the new frame is marked $\tagStackDepth{n + 1}$.

\item From callee back to caller, when execution reaches the corresponding
  return instruction for the call instruction, as defined by
  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
  Prior to this point, the locations of the old frame have been restored to
  $\tagNoDepth$.

@@

These sequences also enforce the $\SP$ discipline. When stack frames are of
fixed size, it suffices to increment $\SP$ and tag it with a dedicated tag in
the header sequence, which is later checked after decrementing $\SP$ during the
exit sequence to authorize the return. The well-formedness of the blessed
sequences can be checked statically. Only programs whose sequences are
well-formed are protected by the micropolicy, although programs are still
protected (by fail-stopping) even if the intended sequences are missing. By
tracking the execution of the blessed sequences in the tag state of the
{\PCname} register, the micropolicy ensures that the entry sequences of calls
declared in the call map are executed correctly.

@@

\paragraph*{Testing the conservative policy}
%
This conservative policy can be shown to satisfy an eager \emph{testing
property}, which at each step of the computation performs integrity and
confidentiality tests for a stack of running variants, which are kept in
lockstep with the original machine.

\definition{Stack integrity test}
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition{Stack confidentiality test}
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\definition{Variant stacks}
  The testing property keeps track of relevant information on a stack whose
  elements represent variants of the machine by recording an initial machine
  state, an initial and current variant of the machine, a contour and a return
  map:
%
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
%
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}

\definition{Variant stack test}
%
  Our testing property extends the basic tests above to whole variant stacks.
  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
  $\mach$ safisfies the \emph{variant stack test} if for each variant element
  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.

@@

  \item If the machine fail-stops, i.e.,
    $(\mach, \pol) \hookrightarrow_P (\mach, \pol, \tau)$,
    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at any
    depth up to the size of $\vs$.

  \item If the top variant of the stack (and only the top variant) returns,
    the property is safisfied for the singleton trace $[(\mach, \pol)]$ at depth
    equal to the size of the stack. \rb{Avoid confusion with return below}

  \item If (1) the machine takes a step,
    $(\mach, \pol) \hookrightarrow_P (\mach', \pol', \obs)$,
    that is neither a call nor a return, (2) all
    variants can also take a step $\vs \hookrightarrow \vs'$ \rb{notation for this}
    while satisfying the eager integrity and confidentiality tests \rb{more formal?}
    (these tests are required also for calls and returns, and omitted in those cases),
    and (3) the post-step state $(\mach', \pol')$ and stack $\vs'$ satisfy the eager testing property
    for a given depth $d$ \rb{omitting stack size condition} and MP-trace $\MPT$,
    then the property is also satisfied for trace $(\mach, \pol) : \MPT$
    at depth $d$.
